Methods for Large Scale Learning to Rank

### 2014-04-29 report/Chapters/LTRMethods.tex
LRUF methode verder uitgewerkt

\section{LRUF}
Learning automaton.
TODO:
\begin{enumerate}
\item Variable structure learning automata uitleggen
\item $L_{R-p}, L_{R-\epsilon p} \text{en} L_{R-I}$ uitleggen
\item $\epsilon$-optimality en absolutely expediency uitleggen en melden dat $L_{R-I}$ deze eigenschappen bezit
\item LRUF-tuple uitleggen
\end{enumerate}

\begin{equation}
p_j(k+1)=	\begin{cases}	p_j(k)+a[1-p_j(k)] 	&\mbox{if } j=i \\ 
							(1-a)p_j(k) 		&\mbox{otherwise} 
			\end{cases} 
\label{eq:lruf_reward_update}
\end{equation}

\begin{equation}
g_i=\frac{1}{N_i}\sum\nolimits_{d_{i}^{j}\in \underline{f}_{i}} a(r_{i}^{j})^{-1}
\label{eq:lruf_average_relevance}
\end{equation}

>>>>

\section{LRUF}
LRUF \cite{Torkestani2012b} is Learning-to-Rank method based on the theory of learning automata. Learning automata are adaptive decision-making units that learn the optimal set of actions through repeated interactions with its environment. Variable structure learning automata can be defined as a triple $<\beta,\alpha,L>$, with $\beta$ being the set of inputs, $\alpha$ the set of actions. $L$, the learning algorithm ,is a recurrence relation that modifies the action probability vector of the actions in $\alpha$.\\

Three well-known learning algorithms $L$ used in variable structure learning automata are \emph{linear reward-penalty} ($L_{R-P}$), \emph{linear reward-$epsilon$-penalty} ($L_{R-\epsilon P}$) and \emph{linear reward-inaction} ($L_{R-I}$). LRUF uses the $L_{R-I}$ learning algorithm, which updates the action probability vector following Equation \ref{eq:lruf_reward_update} when the selected action $a_i(k)$ is rewarded by the environment.
\begin{equation}
p_j(k+1)=	\begin{cases}	p_j(k)+a[1-p_j(k)] 	&\mbox{if } j=i \\ 
							(1-a)p_j(k) 		&\mbox{otherwise} 
			\end{cases} 
\label{eq:lruf_reward_update}
\end{equation}
\noindent where $i$ and $j$ are indices of action in $\alpha$ and $p(k)$ is the probability vector over the action set at rank $k$. $\alpha$ $a$ is the learning rate of the model.\\

A variable action-set learning automaton is an automaton in which the number of available actions change over time. It has been shown that a combination of a variable action-set learning automaton with the $L_{R-I}$ learning algorithm is absolutely expedient and $\epsilon$-optimal \cite{Thathachar1987}, which means that it is guaranteed to approximate the optimal solution to some value $\epsilon$ and each update step is guaranteed not to decrease performance of the model. A variable-action set learning automata has a finite set of $r$ actions $\alpha={\alpha_1,...,\alpha_r}$. $A$ is defined as the power set of $\alpha$, $A={A_1,..,A_m}$ with $m = 2^{r}-1$. $\psi(k)$ is a probability distribution over $A$ such that $\psi(k)=p(A(k)=A_i|A_i\in A, 1\le i \le 2^r-1)$. $\hat{p}_i(k)$ is the probability of choosing action $\alpha_i$ given that action subset $A(k)$ has already been selected and $\alpha_i \in A(k)$.\\

LRUF uses an optimisation problem that can be illustrated as a quintuple $<q_i,A_i,\underline{d_i},\underline{R_i},\underline{f_i}>$, where $q_i$ is a submitted query, $A_i$ is a variable action-set learning automaton, $\underline{d_i}$ is a set of documents associated with $q_i$, and $\underline{R_i}={r_i^j|\forall d_i^j \in \underline{d_i}}$ is a ranking function that assigns rank $r_i^j$ to each document. $\underline{f_i}$ is a feedback set used for the update step described in Equation \ref{eq:lruf_reward_update}. For each rank $k$ the learning automaton $A_i$ chooses one of its actions following its probability vector, jointly forming $\underline{R_i}$. LRUF translates the feedback in $\underline{f_i}$ to an understandable value to use it to converge the action probability vector in optimal configuration. Therefore LRUF defines a $g_i: \underline{f_i} \rightarrow \mathbb{R}^{+}$ to be a mapping from the feedback set into a positive real number. The LRUF mapping function $g_i$ computes the average relevance score of ranking $R_i$ based on $f_i$ and is defined as shown in Equation \ref{eq:lruf_average_relevance}.
\begin{equation}
g_i=\frac{1}{N_i}\sum\nolimits_{d_{i}^{j}\in \underline{f_i}} a(r_{i}^{j})^{-1}
\label{eq:lruf_average_relevance}
\end{equation}
